Formalizing Stack Safety as a Security Property

### 2023-02-03 submission5/core.tex
wibbles in 3
edits, comments in 5

      \}}
  \end{subfigure}
  \begin{subfigure}{\columnwidth}
    \begin{tabular}{r l | l}
      \labeledrow{0:}{addi sp,sp,-20}{\(\mathbf{alloc} ~ (-20,20)\)}
      \labeledrow{4:}{sd ra,12(sp)}{}
      \labeledrow{8:}{sw a0,8(sp)}{}
      \labeledrow{12:}{sw zero,4(sp)}{}
      \labeledrow{16:}{jal f,ra}{\(\mathbf{call} ~ \emplist \)}
      \labeledrow{20:}{sw a0,0(sp)}{}
      \labeledrow{24:}{lw a4,4(sp)}{}
      \labeledrow{28:}{li a5,42}{}
      \labeledrow{32:}{bne a4,a5,L1}{}
      \labeledrow{36:}{lw a0,8(sp)}{}
      \labeledrow{40:}{sw a0,out}{}
      \labeledrow{44:}{j L2:}{}
      \labeledrow{L1, 48:}{lw a0,0(sp)}{}
      \labeledrow{52:}{sw a0,out}{}
      \labeledrow{L2, 56:}{ld ra,12(sp)}{}
      \labeledrow{60:}{addi sp,sp,20}{\(\mathbf{dealloc} ~ (0,20)\)}
      \labeledrow{64:}{jalr ra}{\(\mathbf{return}\)}
    \end{tabular}
  \end{subfigure}
  \begin{subfigure}{\columnwidth}
    \center
    \vspace{\abovedisplayskip}
    \(\textnormal{\large \dots} \pubbox[30px] \!\! \stackrel{\textsc{\large sp}}{\textbox[\tt res]}
    \stackrel{\textsc{\large 4(sp)}}{\textbox[\tt sens]}
    \stackrel{\textsc{\large 8(sp)}}{\textbox[\tt sec]}
    \stackrel{\textsc{\large 12(sp)} \hfill}{\textbox[\(\mathtt{ra_1}\)]\textbox[\(\mathtt{ra_2}\)]}
    \textnormal{\large \dots}\)
  \end{subfigure}

@@

To begin with, what if {\tt f} returns to an unexpected place (i.e. \(\PCname \neq 20\) or
\(\SP \neq 980\))? We consider this to violate \(\wbcf\). \(\wbcf\) is a relationship between
call steps and their corresponding return steps: just after the return, the program
counter should be at the next instruction following the call,
and the stack pointer should be the same as it was before the call.
Both of these are essential. In \cref{subfig:WBCF}, the attacker adds
16 to the return address and then returns, thus bypassing the {\tt if}-test in the code and outputting
{\tt secret}.
In \cref{subfig:WBCF2}, the attacker returns with \(\SP' = 998\) instead of the
correct \(\SP = 980\). In this scenario, given the layout of {\tt main}'s frame,
\begin{center}
\begin{tabular}{| l | l | l | l | l |}
  \multicolumn{1}{r}{\(\SP \downarrow\)} &
  \multicolumn{2}{r}{\(\SP' \downarrow\)} \\
  \hline
  res & sens & sec & ra & ra \\
  \hline
\end{tabular}
\end{center}

@@

Where the caller's confidentiality and integrity are concerned with protecting specific,
identifiable state---the caller's stack frame---their callee equivalents are concerned
with enforcing the expected interface between caller and callee. Communication between
the principals should occur only through the state elements that are designated for the
purpose: those which we label \(\public\) and (in the presence of memory sharing) \(\object\).

Applying this intuition using our framework, \(\clec\) turns out to resemble
\(\clri\), extended to every element that is not marked \(\object\) at call-time. The callee's
internal behavior is represented by those elements that change over the course of its
execution, and which are not part of the interface with the caller. At return, those
elements should become irrelevant.

Similarly, in \(\clei\), only elements marked \(\object\) at the call should influence the
behavior of the callee. This is the same principle as \(\clrc\), extended from \(\sealed\)
elements to all non-\(\object\) ones.

@@

Finally, we can define our core properties. They can be found in \cref{tab:props},
arranged to show their commonalities and distinctions. Each definition gives a criterion
quantified over states \(s\) that immediately follow call (or tail-call) steps.
If an execution can reach a state \(s'\) such that \(s' \stepstounder{\overline{\psi}} s\)
where \(\mathbf{call} ~ \addr ~ \overline{\reg} \in \bar{\psi}\), then \(s\) is the target
of a call.
%Likewise, if \(\mathbf{tailcall} ~ \addr ~ \overline{\reg} \in \bar{\psi}\), then
%\(s\) is the target of a tailcall.
As a shorthand, we write that each property is defined
by a criterion that must hold ``for all call targets \(s\).''

@@

Using this mechanism, an argument passed on the stack via its \(\SP\)-relative offset
uses the operation \(\mathbf{call} ~ \dots ~ (\SP, \mathit{off}, \mathit{sz})\).
C++-style reference passing that places the reference in \(\reg\) uses
\(\mathbf{call} ~ \dots ~ (\reg, 0, \mathit{sz})\).
In the second case, the reference can be passed further down the stack, but otherwise
only the immediate callee gains access.

On the other hand, if the address of an object is taken directly and passed it as a
pointer, we have no way obvious way to limit the scope of such sharing.
We treat this scenario as the caller allocating a ``public''
object that recieves no protection. We extend the \(\mathbf{alloc}\) operation
with a boolean flag, where {\bf t} indicates
that the allocation is public, and {\bf f} that it is private.
In the likely event that the space for multiple objects is allocated in a single step,
that step can make multiple allocation operations, each labeled appropriately.
Public objects are labeled \(\public\) rather than \(\object\), so they are
never sealed at a call (\cref{sfig:publicalloc}).

>>>>

      \}}
  \end{subfigure}
  \begin{subfigure}{\columnwidth}
    \begin{tabular}{r l | l}
      \labeledrow{0:}{addi sp,sp,-20}{\(\mathbf{alloc} ~ (-20,20)\)}
      \labeledrow{4:}{sd ra,12(sp)}{}
      \labeledrow{8:}{sw a0,8(sp)}{}
      \labeledrow{12:}{sw zero,4(sp)}{}
      \labeledrow{16:}{jal f,ra}{\(\mathbf{call} ~ \emplist \)}
      \labeledrow{20:}{sw a0,0(sp)}{}
      \labeledrow{24:}{lw a4,4(sp)}{}
      \labeledrow{28:}{li a5,42}{}
      \labeledrow{32:}{bne a4,a5,L1}{}
      \labeledrow{36:}{lw a0,8(sp)}{}
      \labeledrow{40:}{sw a0,out}{}
      \labeledrow{44:}{j L2:}{}
      \labeledrow{L1, 48:}{lw a0,0(sp)}{}
      \labeledrow{52:}{sw a0,out}{}
      \labeledrow{L2, 56:}{ld ra,12(sp)}{}
      \labeledrow{60:}{addi sp,sp,20}{\(\mathbf{dealloc} ~ (0,20)\)}
      \labeledrow{64:}{jalr ra}{\(\mathbf{return}\)}
    \end{tabular}
  \end{subfigure}
  \begin{subfigure}{\columnwidth}
    \center
    \vspace{\abovedisplayskip}
    \(\textnormal{\large \dots} \pubbox[30px] \!\! \stackrel{\textsc{\normalsize sp}}{\textbox[\tt res]}
    \stackrel{\textsc{\normalsize 4(sp)}}{\textbox[\tt sens]}
    \stackrel{\textsc{\normalsize 8(sp)}}{\textbox[\tt sec]}
    \stackrel{\textsc{\normalsize 12(sp)} \hfill}{\textbox[\(\mathtt{ra_1}\)]\textbox[\(\mathtt{ra_2}\)]}
    \textnormal{\large \dots}\)
  \end{subfigure}

@@

To begin with, what if {\tt f} returns to an unexpected place (i.e. \(\PCname \neq 20\) or
\(\SP \neq 980\))? We consider this to violate \(\wbcf\). \(\wbcf\) is a relationship between
call steps and their corresponding return steps: just after the return, the program
counter should be at the next instruction following the call,
and the stack pointer should be the same as it was before the call.
Both of these are essential. In \cref{subfig:WBCF}, the attacker adds
16 to the return address and then returns, thus bypassing the {\tt if}-test in the code and outputting
{\tt secret}.
In \cref{subfig:WBCF2}, the attacker returns with \(\SP' = 988\) instead of the
correct \(\SP = 980\). In this scenario, given the layout of {\tt main}'s frame,
\begin{center}
\begin{tabular}{| l | l | l | l | l |}
  \multicolumn{1}{r}{\(\SP \downarrow\)} &
  \multicolumn{2}{r}{\(\SP' \downarrow\)} \\
  \hline
  {\tt res} & {\tt sens} & {\tt sec} & \({\mathtt ra_1}\) & \({\mathtt ra_2}\) \\
  \hline
\end{tabular}
\end{center}

@@

Where the caller's confidentiality and integrity are concerned with protecting specific,
identifiable state---the caller's stack frame---their callee equivalents are concerned
with enforcing the expected interface between caller and callee. Communication between
the principals should occur only through the state elements that are designated for the
purpose: those which we label \(\public\) and (in the presence of memory sharing) \(\object\).\apt{Are these backwards?}

Applying this intuition using our framework, \emph{callee confidentiality} (\(\clec\)) turns out to resemble
\(\clri\), extended to every element that is not marked \(\object\) at call-time. The callee's
internal behavior is represented by those elements that change over the course of its
execution, and which are not part of the interface with the caller. At return, those
elements should become irrelevant to the subsequent behavior of the caller.

Similarly, in \emph{callee integrity} (\(\clei\)), only elements marked \(\object\) at the call should influence the
behavior of the callee. This is the same principle as \(\clrc\), extended from \(\sealed\)
elements to all non-\(\object\) ones. \apt{Still not obvious why we're calling this an integrity property.}

@@

Finally, we can define our core properties. They can be found in \cref{tab:props},
arranged to show their commonalities and distinctions. Each definition gives a criterion
quantified over states \(s\) that immediately follow call (or tail-call) steps.\apt{No, we don't
  want to mention tail-calls here. But we do need to explain how the interpretation of the
  property needs to change when we add tail-calls later.}
If an execution can reach a state \(s'\) such that \(s' \stepstounder{\overline{\psi}} s\)
where \(\mathbf{call} ~ \addr ~ \overline{\reg} \in \bar{\psi}\), then \(s\) is the target
of a call.
%Likewise, if \(\mathbf{tailcall} ~ \addr ~ \overline{\reg} \in \bar{\psi}\), then
%\(s\) is the target of a tailcall.
As a shorthand, we write that each property is defined
by a criterion that must hold ``for all call targets \(s\).''

@@

Using this mechanism, an call-by-value argument passed on the stack at an \(\SP\)-relative offset
is specified by the triple \((\SP, \mathit{off}, \mathit{sz})\).
In this case, only the immediate callee gains access to the argument location.
A C++-style call-by-reference argument where the reference is passed in \(\reg\)
is instead specified by \((\reg, 0, \mathit{sz})\), which allows
the reference to be passed further down the stack.
\apt{What if the reference is itself passed on the stack?}
%In this second case, the reference can be passed further down the stack, but otherwise
%only the immediate callee gains access. APT:

On the other hand, if the address of an object is taken directly and passed it as a
pointer, we have no way obvious way to limit the scope of such sharing.
We treat this scenario as the caller allocating a ``public''
object that receives no protection. We extend the \(\mathbf{alloc}\) operation
with a boolean flag, where {\bf t} indicates
that the allocation is public, and {\bf f} that it is private.
If space for multiple objects is allocated in a single step,
that step can make multiple allocation operations, each labeled appropriately.
Public objects are labeled \(\public\) rather than \(\object\), so they are
never sealed at a call (\cref{sfig:publicalloc}).
